Automated reading of residual plot with computer vision models

### 2023-10-03 paper/paper.tex
upload rp training script

---
title: |
  Automated reading of residual plot with computer vision models
type: ARTICLE TEMPLATE
author:
  - name: Weihao Li
    affil: a
    email: weihao.li@monash.edu
affiliation:
  - num: a
    address: |
      Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
# bibliography: ref.bib
abstract: |
  TBD.
keywords: |
  TBD
header-includes: |
  \usepackage{lscape}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
output: rticles::tf_article
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  echo = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = "100%",
  fig.align = "center")
```

```{r}
# OOP supports needed by `visage`
# remotes::install_github("TengMCing/bandicoot")
# 
# Visual inference models and p-value calculation
# remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.0.tar.gz")

library(tidyverse)
library(visage)

# To control the simulation in this file
set.seed(10086)
```

# Introduction

# Methods for training computer vision models to analyze residual plots

@@

## Method 1: Single Plot Model

@@

To implement the Single Plot Model effectively, we utilize established neural network architectures like VGG16. The model is trained on a diverse dataset containing both "not null" plots (generated from model with violations) and "null" plots (generated from model without violations).

@@

1. **Distinguishing null and not null plots**: Through training, the model is expected to be able to distinguish between null plots and not null plots. This distinction forms the foundation for identifying model violations in new residual plots.

2. **Capturing variance in null plots**: Null plots often exhibit varying shapes and outliers due to the randomness of the error term and the distribution of fitted values. The model learns to recognize and characterize this variance, aiding in the detection of violations.

3. **Recognizing patterns in not null plots**: The model would develop the ability to identify common patterns in not null plots generated through simulations.

4. **Analyzing Aesthetic Elements**: Beyond content, the model takes into account aesthetic elements within the residual plot, including point size, auxiliary lines, layers, background colors, and other visual attributes. This is an undesired capability but it is unavoidable if only a single design of plot is used as inputs.

@@

improvements:
1. enrich the factors in simulation
2. sigma same, shape same, same fitted value distribution, stack residual plots  
3. test other residual plots generated from factors not seen by the model
4. replace the output with the effect size

@@

## Lineup model

@@

1. **Locating the data plot**: It learns to identify and locate the data plot within a lineup of residual plots.

2. **Comparing data plot and null plots**: The model may potentially compare the data plot with null plots within the lineup. This provide additional information for decision making.

3. **Understanding variance within a lineup**: It may also understand the variance among null plots within a lineup.

@@

## Position model

@@

## Effect size model

##

>>>>

% interactcadsample.tex
% v1.03 - April 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{lscape}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\usepackage{setspace}

\begin{document}

\articletype{ARTICLE TEMPLATE}

\title{Automated reading of residual plot with computer vision models}

\author{\name{Weihao Li$^{a}$}
\affil{$^{a}$Department of Econometrics and Business Statistics, Monash
University, Clayton, VIC, Australia}
}

\thanks{CONTACT Weihao
Li. Email: \href{mailto:weihao.li@monash.edu}{\nolinkurl{weihao.li@monash.edu}}}

\maketitle

\begin{abstract}
TBD.
\end{abstract}

\begin{keywords}
TBD
\end{keywords}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{methods-for-training-computer-vision-models-to-analyze-residual-plots}{%
\section{Methods for training computer vision models to analyze residual
plots}\label{methods-for-training-computer-vision-models-to-analyze-residual-plots}}

@@

\hypertarget{method-1-single-plot-model}{%
\subsection{Method 1: Single Plot
Model}\label{method-1-single-plot-model}}

@@

To implement the Single Plot Model effectively, we utilize established
neural network architectures like VGG16. The model is trained on a
diverse dataset containing both ``not null'' plots (generated from model
with violations) and ``null'' plots (generated from model without
violations).

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Distinguishing null and not null plots}: Through training, the
  model is expected to be able to distinguish between null plots and not
  null plots. This distinction forms the foundation for identifying
  model violations in new residual plots.
\item
  \textbf{Capturing variance in null plots}: Null plots often exhibit
  varying shapes and outliers due to the randomness of the error term
  and the distribution of fitted values. The model learns to recognize
  and characterize this variance, aiding in the detection of violations.
\item
  \textbf{Recognizing patterns in not null plots}: The model would
  develop the ability to identify common patterns in not null plots
  generated through simulations.
\item
  \textbf{Analyzing Aesthetic Elements}: Beyond content, the model takes
  into account aesthetic elements within the residual plot, including
  point size, auxiliary lines, layers, background colors, and other
  visual attributes. This is an undesired capability but it is
  unavoidable if only a single design of plot is used as inputs.
\end{enumerate}

@@

improvements: 1. enrich the factors in simulation 2. sigma same, shape
same, same fitted value distribution, stack residual plots\\
3. test other residual plots generated from factors not seen by the
model 4. replace the output with the effect size

@@

\hypertarget{lineup-model}{%
\subsection{Lineup model}\label{lineup-model}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Locating the data plot}: It learns to identify and locate the
  data plot within a lineup of residual plots.
\item
  \textbf{Comparing data plot and null plots}: The model may potentially
  compare the data plot with null plots within the lineup. This provide
  additional information for decision making.
\item
  \textbf{Understanding variance within a lineup}: It may also
  understand the variance among null plots within a lineup.
\end{enumerate}

@@

\hypertarget{position-model}{%
\subsection{Position model}\label{position-model}}

@@

\hypertarget{effect-size-model}{%
\subsection{Effect size model}\label{effect-size-model}}

\end{document}
